General constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-24 magic_paper.tex
Edits from pdf implemented. Few small things remaining for tomorrow.

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g. free energy changes in the case of temperature, or non-unitality in the case of the maximally mixed state). Such bounds are therefore of interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in a few ways. We obtain distillation upper bounds without the use of monotones, but instead tools from statistical mechanics and recent work from the overlap of thermodynamics and quantum information theory~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool in thermodynamics and information theory -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore our analysis is also of interest to the study of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

\subsection{Overview of our results}
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations (Section~\ref{sec:ps}). We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels (Section~\ref{sec:unital}). These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed point (Section~\ref{sec:frag}). We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that lead to a simple description in the thermodynamic limit (Section~\ref{sec:stab}).

@@

We find that the analysis in the presence of negativity is non-trivial, displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by Lorenz curves having a non-trivial maximum related to mana (Section~\ref{sec:lc}) and by showing that single-shot entropies for general Wigner quasi-distributions are well-defined, obey data-processing under stabilizer operations and can take on negative values (Section~\ref{section:entropies}).

@@

We also show that the mana of a quantum state arises as a residue in the $\alpha$--parameter for Renyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Equation (\ref{eq:mana-divergence})).
Finally, in Section~\ref{sec:lower_bounds} we discuss how our approach could be improved and how it can be exploited to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d := \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d = d_1 \dots d_n$ we can decompose the Hilbert space as $\cal{H}_d = \cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, and then define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n} \eqqcolon  \cal{P}_d,
\end{align*}
to denote the phase space point for the composite system. 
For simplicity going forward, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmz}: k \in \mathbb{Z}_d, \bmz \in \cal{P}_d^n\}.
\end{equation}
The Clifford operations $ \cal{C}_d^n $ are then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabiliser states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3} and, finally, the full set of stabilizer states as the convex hull of all pure stabilisers, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is Clifford.

@@

In order to provide a complete decomposition of arbitrary quantum states and quantum operations we now define a complete basis of Hermitian observables that behaves naturally under the action of the Clifford group. At every point $\bmz \in \P_d$ we define the phase-point operator
\begin{equation}\label{eq:ax}
	A_{\bmz} \coloneqq \frac{1}{d} \sum_{\bmy \in \cal{P}_d} \omega^{\eta(\bmz, \bmy)} D_{\bmy}, 
\end{equation}
where $\eta(\bmz, \bmz)$ is the symplectic inner product between any two points $\bmz,\z \in \P_d$, and is given explicitly by
\begin{equation}
	\eta(\bmz, \bmy) \coloneqq \bmy^T \begin{pmatrix}
		0  & \id \\ %\mathbbm{O}_n  & \id_n \\
		-\id & 0 \\ %-\id_n & \mathbbm{O}_n \\
	\end{pmatrix} \bmz,
\end{equation}
where $0, \id$ denote the $n\times n$ zero and identity matrices.

@@

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmz)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\bmz)$ is a bounded real-valued function on $\P_d$ with the normalisation property $\sum_{\bmz} W_\rho(\bmz) = 1$ (see~\cref{app:wigner} for details).

Any quantum operation $\E$ can also admit a Wigner representation. If $\E$ maps some quantum system $A$ of dimension $d_A$ to a quantum system $B$ of dimension $d_B$ and $\J(\E)$ is its associated Choi state~\cite{watrous_2018}, then we can define
\begin{equation}
W_{\E}(\y |\bmz) \coloneqq d_A^2 W_{\J(\E)}(\bar{\bmz} \oplus \y),
\end{equation}
where $\bar{\bmz} = (q_1, \dots, q_n, -p_1, \dots, -p_n)$ can be viewed as the time-reversed version of $\bmz$ in the discrete phase space, where momenta are reflected while position coordinates remain unchanged.

@@

In any theory of magic, one route to bounding distillation rates is through a \emph{magic monotone}. A magic monotone is a real-valued function of any quantum state $\M(\rho)$ that is monotonically non-increasing under the free operations of the magic theory. More precisely $\M(\sigma) \le \M(\rho)$ whenever it is possible to convert $\rho$ into $\sigma$ using free operations.

One prominent magic monotone is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \mana{\rho} \coloneqq \ln (2\hspace{1pt}\sn{\rho}+1),
\end{equation}
where the \emph{sum-negativity}~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmz: \W[\bmz]{\rho} < 0} \abs{\W[\bmz]{\rho}}.
\end{equation}

Mana is an additive\footnote{It satisfies the condition $\mana{\rho_1 \otimes \rho_2} = \mana{\rho_1} + \mana{\rho_2}$.} magic monotone, and the fact that it is non-increasing under free operations provides a constraint on magic state interconvertibility.

In this work we shall develop bounds that apply to essentially any reasonable magic theory. The central idea is to apply majorization theory to the quasi-probability distributions that arise for magic states. We begin in the next section by explaining how majorization relates to the free operations in a theory of magic.

@@

Within the Wigner representation for odd prime dimension, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:veitch}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory are assumed to lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmz]{\rho} \geq 0 \text{ for all } \bmz \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed quantum operations, i.e. the ones that are considered free within the theory. The most basic assumption we require on free operations is that they send any free state to another free state, meaning that a resource state cannot be created freely, a common assumption among all magic theories.

Any magic state protocol will correspond to a quantum channel $\E$, and admits a Wigner representation $W_{\E}( \y | \bmz)$ that can be viewed as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\E$ and $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

However, it is important to note that this mapping to stochastic maps is not one-to-one. Not all stochastic maps on the phase space correspond to valid quantum operations. The reason is that the maps must also respect the symplectic structure of the phase space, which is an additional non-trivial constraint.

In what follows, we shall assume that we have a magic theory $\R = (\F, \O)$ in which the free states $\F$ form a closed set and are represented by non-negative Wigner functions, while the free operations $\O$ are stochastic maps in the Wigner representation. To aid our analysis of magic in this phase space setting we shall make use of majorization theory, which we describe in the next section.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has recently found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations. In its simplest form it defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. A well-known result~\cite{cit:marshall} tells us that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

There is a natural generalisation, which is called $\bmd$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization~\cite{cit:horodecki2013}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as $\p \succ_{\r} \q$, if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, this majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from $\bmd$-majorization is also easily checked and again corresponds to checking a small number of inequalities.

In fact, we can further generalise to \emph{relative majorization}~\cite{Blackwell_1953, Ruch_1976, ruch_mixing_1978, Renes_2016, Buscemi_2017, Rethinasamy_2020}, defined as an ordering between pairs of vectors so that 
\begin{equation}
	(\p,\r) \succ (\q, \r')
\end{equation}
iff there is a stochastic map $A$ such that $A\r = \r'$ and $A\p = \q$. We retrieve $\bmd$--majorization when $\r = \r'$.

It turns out that relative majorization is equally applicable to \emph{quasi-probability} distributions, and therefore can be applied to the Wigner representation of magic across different $\R_\sigma$. We discuss this in detail in the next section.

@@

Central to the analysis is the notion of a Lorenz curve of a vector $\w \in \mathbb{R}^n$ relative to some other vector $\r \in \mathbb{R}^n$. Given a vector $\w$ we define $\w^\downarrow$ to be the re-arrangement of the components of $\w$ into decreasing order. Given two $n$--component vectors $\w$ and $\r$, we first define $\widetilde{\w}=(\widetilde{w}_i)$, where $\widetilde{w}_i \coloneqq w_i/r_i$, as the vector of component-wise ratios between $\w$ and $\r$.
We can now define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}
\label{eq:lc}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R\coloneqq \sum_{i=1}^n r_i$ and $\pi$ is the permutation on $n$ objects mapping $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as the rank of a quantum state is not operationally meaningful.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotonically increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See~\cref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality. Since majorization provides a measure of order/disorder this is consistent with the intuition that negativity, for example within the context of quantum optics or quantum computing, can provide a greater form of order than is possible within classical theory.

@@

However, relative majorization is usually considered for probability distributions, so we need to verify that the same Lorenz curve conditions apply to quasi-probability distributions before proceeding with our analysis. This can be done from first principles, but a simpler way to see that this holds is to reduce to the problem of genuine distributions by first `masking' the negativity in the quasi-probability distribution $\w$ using the reference distribution $\r$ and then applying the conditions for relative majorization. This negativity masking then gives the following result.

@@

We can now consider how physical constraints affect magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This allows the use of $\mathbf{d}$--majorization, however in~\cref{sec:stab}, we drop the requirement of fixed-points and instead use the full relative majorization, before finally giving general entropic constraints in Section~\ref{section:entropies}.

>>>>

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated in the distillation bounds. This is achieved by considering how a given magic state protocol transforms a \emph{pair} of quantum states -- one being a noisy magic state, and the other a stabilizer state that is distinguished by the physics of the system (for example a state at a characteristic temperature, or the maximally mixed state) and acts as a reference state for the protocol. The magic distillation bounds can then be expressed in terms of the physics of the reference state (e.g.~free energy changes in the case of temperature). Such bounds are therefore of interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols~\cite{Tuckett_2019, Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in key ways. We obtain distillation upper bounds without the use of monotones, but instead use tools from statistical mechanics and recent work in single-shot resource theories~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. Our analysis relies on the discrete Wigner representation of quantum systems, in which all states and operations can be described on a discrete phase space~\cite{Ferrie_2008, Okay_2021}. Crucially, we focus on magic states with negativity in their Wigner representation, which is known to be a necessary condition for universality in the state-injection model~\cite{cit:veitch, cit:mari, cit:gottesman, cit:knill, Campbell_2011}. However, taking a statistical mechanical perspective in this context raises a problem: the standard Boltzmann entropy is not defined for quasi-distributions. We circumvent this obstacle by making use of a more fundamental tool -- majorization theory~\cite{cit:marshall, Veinott_1971, Ruch_1976}. To our knowledge majorization of quasi-distributions has not been considered in quantum physics before, and therefore our analysis is also of relevance to other studies of non-classicality in quantum systems~\cite{Fine_1982, Allahverdyan_2018, arvidsson_2020, halpern_2018, Lostaglio_2018, Levy_2020}.

\subsection{Overview of our results}
We begin by describing how stabilizer operations correspond to stochastic maps on a discrete phase space, and how majorization can be applied to constrain magic state transformations (Section~\ref{sec:ps}). We then consider families of magic protocols with increasing complexity. The simplest case we consider is for magic protocols that lead to unital channels (Section~\ref{sec:unital}). These are a subset of protocols that leave invariant some distinguished state or have some equilibrium fixed-point structure (Section~\ref{sec:frag}). We then consider protocols that are non-equilibrium processes, but only generate sub-linear correlations that enable a simple description in the thermodynamic limit (Section~\ref{sec:stab}), before giving entropic analysis that can be applied more generally (Section~\ref{section:entropies}).

@@

We find that the analysis in the presence of negativity is non-trivial, displays a range of features that do not appear in classical statistical mechanics, and leads to a picture of Wigner negativity in a quantum circuit being described as non-classical free energy that is processed under stochastic dynamics. This is demonstrated by Lorenz curves having a non-trivial maximum that is related to the mana of a state (Section~\ref{sec:lc}), and by showing that single-shot entropies for general Wigner quasi-distributions are well-defined, obey data-processing under stabilizer operations and can take on negative values (Section~\ref{section:entropies}).

@@

We show that the mana of a quantum state also arises as a residue in the $\alpha$--parameter for Renyi entropies, and gives the rate of divergence to $-\infty$ as we approach the Shannon entropy of a Wigner distribution (Equation (\ref{eq:mana-divergence})).
Finally, in Section~\ref{sec:lower_bounds} we discuss how our approach could be extended and how it can be exploited to construct explicit lower bounds on distillation that take into account physical constraints.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d := \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmz \coloneqq (q, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmz} \coloneqq \tau^{q p} X^{q} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime~\cite{Appleby_2005}. The qubit case $d=2$ is known to have obstacles to using a Wigner representation to quantify magic in terms of negativity~\cite{Mermin_1990, cit:howard2}, although recent work~\cite{Raussendorf_qubits} has shown how to extend this to include qubits by enlarging the phase space. For a composite system with dimension $d_{\rm tot} = d_1 \dots d_n$ and Hilbert space $\cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, we define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmz} \coloneqq D_{(q_1, p_1)} \otimes \dots \otimes D_{(q_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmz &\coloneqq (q_1, \dots, q_n, p_1, \dots, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n},
\end{align*}
to denote the phase space point for the composite system. 
For simplicity going forward, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\bmz \in \mathbb{Z}_d^{2n}$.

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmz}: k \in \mathbb{Z}_d, \bmz \in \cal{P}_d^n\}.
\end{equation}
The Clifford group $ \cal{C}_d^n $ is then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabiliser states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3} and, finally, the full set of stabilizer states as the convex hull of all pure stabilisers, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is Clifford.

@@

In order to provide a complete decomposition of arbitrary quantum states and quantum operations we now define a basis of Hermitian observables that transforms covariantly under the action of the Clifford group. At every point $\bmz \in \P_d$ we define the phase-point operator
\begin{equation}\label{eq:ax}
	A_{\bmz} \coloneqq \frac{1}{d} \sum_{\bmy \in \cal{P}_d} \omega^{\eta(\bmz, \bmy)} D_{\bmy}, 
\end{equation}
where $\eta(\bmz, \bmz)$ is the symplectic inner product between any two points $\bmz,\z \in \P_d$, and is given explicitly by
\begin{equation}
	\eta(\bmz, \bmy) \coloneqq \bmy^T \begin{pmatrix}
		0  & \id \\ %\mathbbm{O}_n  & \id_n \\
		-\id & 0 \\ %-\id_n & \mathbbm{O}_n \\
	\end{pmatrix} \bmz,
\end{equation}
where $0, \id$ denote the $n\times n$ zero and identity matrices.

@@

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmz)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\bmz)$ is a real-valued function on $\P_d$ with the normalisation property $\sum_{\bmz} W_\rho(\bmz) = 1$ (see~\cref{app:wigner} for details).

Any quantum channel $\E$ also admits a Wigner representation. If $\E$ maps some quantum system $A$ of dimension $d_A$ to a quantum system $B$ of dimension $d_B$ and $\J(\E) = \E \otimes id (|\phi^+\> \<\phi^+|)$, with $|\phi^+\> = d_A^{-1/2} \sum_k |kk\>$, is its associated Choi state of the channel~\cite{watrous_2018}, then we can define
\begin{equation}\label{Wigner-Choi}
W_{\E}(\y |\bmz) \coloneqq d_A^2 W_{\J(\E)}(\bar{\bmz} \oplus \y),
\end{equation}
where $\bar{\bmz} = (q_1, \dots, q_n, -p_1, \dots, -p_n)$ can be viewed as the time-reversed version of $\bmz$ in the discrete phase space, where momenta are reflected while position coordinates remain unchanged.

@@

In any theory of magic, one route to bounding distillation rates is through a \emph{magic monotone}. A magic monotone is a real-valued function of any quantum state $M(\rho)$ that is monotonically non-increasing under the free operations of the magic theory. More precisely $M(\sigma) \le M(\rho)$ whenever it is possible to convert $\rho$ into $\sigma$ using free operations.

One prominent magic monotone is the \emph{mana} of a state~\cite{cit:veitch2}, defined as
\begin{equation}
    \M(\rho) \coloneqq \log (2\hspace{1pt}\sn{\rho}+1),
\end{equation}
where the \emph{sum-negativity}~\cite{cit:veitch2} is the sum of the negative components in $\W{\rho}$,
\begin{equation}
    \sn{\rho} \coloneqq \sum\limits_{\bmz: \W[\bmz]{\rho} < 0} \abs{\W[\bmz]{\rho}}.
\end{equation}
Using the fact that $W_\rho(\z)$ is a normalised quasi-distribution, we can also write $\M(\rho) = \sum_{\z} |W_\rho(\z)|$.

Mana is an additive magic monotone, and the fact that it is non-increasing under free operations provides a constraint on magic state interconvertibility.

In this work we develop bounds that apply to any reasonable magic theory. The central idea is to apply majorization theory to the quasi-probability distributions that arise for magic states. We begin in the next section by explaining how majorization relates to the free operations in a theory of magic.

@@

Within the Wigner representation for odd prime dimension, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:veitch}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory are required to be a subset of
\begin{equation}
    \F \coloneqq \{ \rho: \W[\bmz]{\rho} \geq 0 \text{ for all } \bmz \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed quantum operations, i.e. the ones that are considered free within the theory. The most basic assumption we require on free operations is that they send any free state to another free state, meaning that a resource state cannot be created freely.

Any magic state protocol will correspond to a quantum channel $\E$, and so from Equation~(\ref{Wigner-Choi}) admits a Wigner representation $W_{\E}( \y | \bmz)$ that acts as a transition matrix mapping phase space points $\bmz \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\bmz \in \P_d} W_{\E}( \y | \bmz) W_\rho(\bmz),
\end{equation}
for any $\E$ and $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that if $\E$ is a free operation then the associated Wigner representation $W_{\E}(\y |\bmz)$ must form a stochastic matrix. In particular, all stabilizer operations correspond to stochastic matrices in the Wigner representation.

However, note that this mapping to stochastic maps is not one-to-one. Not all stochastic maps on the phase space correspond to valid quantum operations. The reason is that the maps must also respect the symplectic structure of the phase space, which is an additional non-trivial constraint.

In what follows, we shall assume that we have a magic theory $\R = (\F, \O)$ in which the free states $\F$ are represented by non-negative Wigner functions, while the free operations $\O$ are stochastic maps in the Wigner representation. Our analysis of magic in this phase space setting makes use of majorization theory, which we describe in the next section.

@@

Majorisation~\cite{cit:marshall, Blackwell_1953} is a collection of powerful tools that has found many applications in quantum information theory~\cite{Nielsen_1999, cit:cwiklinski, cit:lostaglio2, cit:gour, cit:gour2, Horodecki_2003, Vallejos_2021}.
It describes the disorder of distributions that undergo stochastic transformations, and in its simplest form defines a pre-order on probability distributions. Given two distributions $\p= (p_1, \dots, p_n)$ and $\q = (q_1, \dots, q_n)$ over $n$ outcomes, we say that $\p$ majorises $\q$, denoted $\p \succ \q$, if there exists a bistochastic map $A = (A_{ij})$ such that $A\p = \q$, where bistochastic means that $A_{ij} \geq 0$ and $\sum_i A_{ij} = \sum_j A_{ij} = 1$. A well-known result~\cite{cit:marshall} tells us that the condition $ \p \succ \q$ over probability distributions is equivalent to $n-1$ inequalities, which can be checked efficiently.

There is a natural generalisation, which is called $\bmd$--majorization~\cite{Veinott_1971}, or in the context of thermodynamics, thermo-majorization~\cite{cit:horodecki2013}. For a fixed probability distribution $\r = (r_1, \dots, r_n)$ with positive components, we define majorization relative to $\r$ as $\p \succ_{\r} \q$, if and only if there exists a stochastic map $A$ such that $A\r = \r$ and $A \p = \q$. In order words, this majorization pre-order coincides with the image of $\p$ under stochastic maps that have $\r$ as a fixed point. It is readily checked that the original majorization condition between probability distributions corresponds to the case $\r = (1/n, \dots, 1/n)$. The pre-order that results from $\bmd$-majorization is also easily computed and again corresponds to checking a small number of inequalities.

In fact, we can further generalise to \emph{relative majorization}~\cite{Blackwell_1953, Ruch_1976, ruch_mixing_1978, Renes_2016, Buscemi_2017, Rethinasamy_2020}, defined as an ordering between pairs of vectors and write 
\begin{equation}
	(\p,\r) \succ (\q, \r')
\end{equation}
if and only if there is a stochastic map $A$ such that $A\r = \r'$ and $A\p = \q$. We retrieve $\bmd$--majorization when $\r = \r'$.

@@

We also make use of the notion of a Lorenz curve of a vector $\w \in \mathbb{R}^n$ relative to some other vector $\r \in \mathbb{R}^n$. Given a vector $\w$ we define $\w^\downarrow$ to be the re-arrangement of the components of $\w$ into decreasing order. Given two $n$--component vectors $\w$ and $\r$, we first define $\widetilde{\w}=(\widetilde{w}_i)$, where $\widetilde{w}_i \coloneqq w_i/r_i$, as the vector of component-wise ratios between $\w$ and $\r$.
We can now define the \emph{Lorenz curve} of $\w$ relative to $\r$, denoted $L_{\w|\r}(x)$, as the piece-wise linear function that passes through $(0,0)$ and the $n$ points
\begin{equation}
\label{eq:lc}
        (x_k,L_{\w|\r}(x_k)) =\left( \frac{1}{R}\sum_{i=1}^k r_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right),
\end{equation}
where $R\coloneqq \sum_{i=1}^n r_i$ and $\pi$ is the permutation on $n$ objects mapping $\widetilde{\w}$ to $\widetilde{\w}^{\downarrow}$. The form of this requires that $\r$ has no zero components, which we shall assume without loss of generality as the rank of a quantum state is not operationally meaningful.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotonically increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See~\cref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality.

@@

Relative majorization is usually considered for probability distributions, so we need to verify that the same Lorenz curve conditions apply to quasi-probability distributions before proceeding with our analysis. This can be done from first principles, but a simpler way to see that this holds is to reduce to the problem of genuine distributions by first `masking' the negativity in the quasi-probability distribution $\w$ using the reference distribution $\r$ and then applying the conditions for relative majorization. This negativity masking then gives the following result.

@@

We can now consider how physical constraints affect magic distillation rates starting in this section with an abstract setting based on the fixed-point structure of the physical operations accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This allows the use of $\mathbf{d}$--majorization, however in~\cref{sec:stab}, we drop the requirement of fixed-points and instead use the full relative majorization, before giving general entropic constraints in Section~\ref{section:entropies}.
